
TITLE  Exploratory data Analysis to determine the quality of red wine.
Author - VINAYAK NAYAK
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(gridExtra)
library('bitops')
library('RCurl')
library(GGally)
library(scales)
library(memisc)
```

```{r echo=FALSE}
# Load the Data
setwd('C:/Users/vnayak/Documents/vnayak_data_bkp/Udacity/DataAnalyst_Nano/Data_Analysis_R/FINAL_PROJECT/FINAL_PROJ_SUBMISSION/github_R')

getwd()
#list.files()
redwinedf<-read.csv('wineQualityReds.csv')
names(redwinedf)
str(redwinedf)
summary(redwinedf)
```

Our data set contains 1599 observations of 13 variables.


# Univariate Plots Section
```{r echo=FALSE}
# We will analyze the quality variable.
ggplot(aes(x=quality),data = redwinedf)+
  geom_histogram(binwidth = 0.5)+
  scale_x_continuous(breaks = seq(3,8,0.5))

```

```{r Analysis of variable quality}
```

```{r echo=FALSE}
summary(redwinedf$alcohol)
ggplot(aes(x=alcohol),data = redwinedf)+
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(8.40,14.90,0.5))
```
 In the above graph we see that the data is skewed to the right. So we will apply log #transformation to the x axis. We see bi modal distribution in our graph.

```{r echo=FALSE}
ggplot(aes(x=alcohol),data = redwinedf)+
  geom_histogram(binwidth = 0.01)+
  scale_x_log10()
```

```{r echo=FALSE }
ggplot(aes(x=quality),data = subset(redwinedf,alcohol>=9.5))+
  geom_histogram(binwidth = 0.5)
```



```{r echo=FALSE}
#Analyze the sulphate variable
summary(redwinedf$sulphates)
ggplot(aes(x=sulphates),data = redwinedf)+
  geom_histogram(binwidth = 0.05)+
  scale_x_continuous(breaks = seq(0.3,2,0.1))



```

```{r echo=FALSE}
# Analyzed the distribution of the variable. Its normally distributed. 
ggplot(aes(x=sulphates),data = redwinedf)+
  geom_histogram(binwidth = 0.01)+
  scale_x_log10(breaks=seq(0.3,1,0.1))

```


```{r echo=FALSE}
ggplot(aes(x=pH),data = redwinedf)+
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(2.740,4.010,0.1))
```




```{r echo=FALSE}
ggplot(aes(x=density),data = redwinedf)+
  geom_histogram(binwidth = 0.001)+
  scale_x_continuous(breaks = seq(0.990,1.0037,0.001))
```



```{r echo=FALSE}
summary(redwinedf$residual.sugar)
ggplot(aes(x=residual.sugar),data = redwinedf)+
  geom_histogram(binwidth = 0.5)+
  scale_x_continuous(breaks = seq(0.900,15,1))
```

The graph above shows it is skewed to right. We will apply log transformation.
```{r echo=FALSE}
ggplot(aes(x=residual.sugar),data = redwinedf)+
  geom_histogram(binwidth = 0.1)+
  scale_x_log10(breaks=seq(1,10,1))
```


```{r echo=FALSE}
ggplot(aes(x=residual.sugar),data = subset(redwinedf,redwinedf$residual.sugar <2.6))+
  geom_histogram(binwidth = 0.08)+
  scale_x_continuous(breaks = seq(0.90,2.6,0.1))
```

I will create a new variable called quality.factor which will be helpful in plotting box plots . Since quality it is a continous variable we will make the new variable as factor var. 
```{r echo=FALSE}
redwinedf$quality.factor<-cut(redwinedf$quality,c(2,4,5,6,7,8,9))
```





# Univariate Analysis

Structure of the dataset.
Ans:  There are 1599 observation with 13 variables (x,fixed.acidity,volatile.acidity,citric.acid,residual.sugar,chlorides,free.sulphur.dioxide,density,pH,sulphates,alcohol,quality).X and quality are int variable and rest of the variables are num.

#Other observations:
quality has minimum value of 3. Median is 6. Mean is 5.635, MAx is 8.
The count of density  is more with density value from 0.996 to 0.997
residual sugar has median value as 2.200,mean is 2.539.  But the max value is 15.500 which seems to be an outlier in the data.
fixed.acidity has median value as 7.90,mean as 8.32, but the max value os 15.90 which is an outlier.



#Main feature(s) of interest in dataset
The  main feature of interest in your dataset is  to observe the quality of the red wine.Also how other factors like alcohol,pH,residual sugar affect the quality of the red wine.


#Other features in the dataset I think will help support my investigation into the feature(s) of interest?
Fixed.acidity,volatile.acidity and citric.acid are other features that may help in #investigation

#Create any new variables from existing variables in the dataset
New variable created is ::  redwinedf$quality.factor

#Unusual distributions.Operation performed to change distribution.
Alcohol was skewed to the right. So I applied lox_x transformation and observed a bimodal distribution.
residual sugar was skewed to the right. After applying log_x transformation it changed into a normal distribution.



# Bivariate Plots Section
```{r echo=FALSE}
p1<-ggplot(aes(x=quality,y=alcohol),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r correlation between quality and alcohol,echo=FALSE}
cor.test(redwinedf$quality,redwinedf$alcohol)

```

```{r scatter plot for quality and pH,echo=FALSE}
p2<-ggplot(aes(x=quality,y=pH),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r correlation between quality and pH,echo=FALSE}
cor.test(redwinedf$quality,redwinedf$pH)
```

```{r scatterplot for quality vs residual sugar,echo=FALSE}
p3<-ggplot(aes(x=quality,y=residual.sugar),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r correlation,echo=FALSE}
cor.test(redwinedf$quality,redwinedf$residual.sugar)
```

```{r fixed.acidity and quality,echo=FALSE}
p4<-ggplot(aes(x=quality,y=fixed.acidity),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r correlation between quality and fixed.acidity}
cor.test(redwinedf$quality,redwinedf$fixed.acidity)
```

```{r quality and volatile.acidity}
p5<-ggplot(aes(x=quality,y=volatile.acidity),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r correlation between quality and volatile.acidity}
cor.test(redwinedf$quality,redwinedf$volatile.acidity)
cor.test(redwinedf$quality,redwinedf$citric.acid)
```

```{r quality and citric acid}
p6<-ggplot(aes(x=quality,y=citric.acid),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```


```{r quality and chlorides}
p7<-ggplot(aes(x=quality,y=chlorides),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r quality and free.sulphur.dioxide}
p8<-ggplot(aes(x=quality,y=free.sulfur.dioxide),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r quality and total sulphur dioxide}
p9<-ggplot(aes(x=quality,y=total.sulfur.dioxide),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r quality and sulphates}
p10<-ggplot(aes(x=quality,y=sulphates),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```

```{r quality and density}
p11<-ggplot(aes(x=quality,y=density),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')
```


```{r}
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,ncol=3)
```


```{r relationship between fixed.acidity and volatile.acidity}
ggplot(aes(x=fixed.acidity,y=volatile.acidity),data = redwinedf)+
  geom_point()+
  geom_smooth(method = 'lm',color='red')

cor.test(redwinedf$fixed.acidity,redwinedf$volatile.acidity)
```

```{r relationship between free.sulphur.dioxide and total.sulfur.dioxide}
ggplot(aes(x=free.sulfur.dioxide,y=total.sulfur.dioxide),data = redwinedf)+
  geom_point(alpha=1/10,position = position_jitter(h=0))+
  geom_smooth(method = 'lm',color='red')

cor.test(redwinedf$free.sulfur.dioxide,redwinedf$total.sulfur.dioxide)
```
 We see that there is strong relationship between free.sulphur.dioxide and #total.sulphur.dioxide.


```{r box plot between quality.factor and alcohol}
ggplot(aes(x=quality.factor,y=alcohol),data = redwinedf)+
  geom_boxplot()

```

```{r box plot between quality.factor and citric.acid}
ggplot(aes(x=quality.factor,y=citric.acid),data = redwinedf)+
  geom_boxplot()
```

```{r box plot between quality.factor and sulphates}
ggplot(aes(x=quality.factor,y=sulphates),data = redwinedf)+
  geom_boxplot()
```




# Bivariate Analysis


 Alcohol,citric acid and suplhates are strongly co related to quality. Fixed acidity is slightly co related to quality. It can be seen from the above graph.




#Interesting relationships between the other features (not the main feature(s) of interest)?
 We see that there is strong relationship between free.sulphur.dioxide and total.sulphur.dioxide.

# Strongest relationship.
Strongest correlation between free.sulfur.dioxide and total.sulfur.dioxide.
Second co relation is beween alcohol and quality. 



# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

```{r Scatterplot with quality and alcohol using color}
ggplot(aes(x=quality,y=alcohol,color=quality.factor),data = redwinedf)+
  geom_point(alpha=1/5,position = position_jitter(h=0))
```
It is clear from the above graph that the quality is determined by the  content of alcohol 

```{r}
ggplot(aes(x=quality,y=citric.acid,color=quality.factor),data = redwinedf)+
  geom_point(alpha=1/5,position = position_jitter(h=0))
```

```{r}

ggplot(aes(x=quality,y=sulphates,color=quality.factor),data = redwinedf)+
  geom_point(alpha=1/2,position = position_jitter(h=0))+
  ylim(0,quantile(redwinedf$sulphates,0.99))+
  scale_color_brewer(type = 'qual')

```
 Alcohol, citric acid and Sulphate are strongly related to quality in a linear way.As the value of the three variables increases the quality of the alcohol increases.


```{r}
ggplot(aes(x=quality,y=total.sulfur.dioxide,color=quality.factor),data = redwinedf)+
  geom_point(alpha=1/5,position = position_jitter(h=0))
```

```{r quality and volatile acidity, non linear relationship}

ggplot(aes(x=quality,y=volatile.acidity,color=quality.factor),data = redwinedf)+
  geom_point(alpha=1/5,position = position_jitter(h=0))+
  scale_color_brewer(type = 'qual')
```
 From the above observation we see that quality of the alcohol increases with the decrease in volatile.acidity.So there is non linear relationship.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
#Relationship between quality and alcohol shows strong relationship between the two variables.

alcohol - Strong linear
pH - Slight Non linear
residual sugar - flat
fixed.acidity - slight linear
volatile.acidity - Strong non linear
citric.acid - Strong linear
chlorides - Slight non linear (close to flat)
free.sulfur.dioxide - Slight non linear (close to flat)
total sulphur dioxide - Non linear
Sulphates - Strong linear
density - Non linear


### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
